\documentclass{note}
%\addtolength{\topmargin}{-2cm}
%\addtolength{\textheight}{2cm}
\usepackage{mathptm,mydef}
\usepackage{courier}
\usepackage{alltt}
%\renewcommand{\ttdefault}{txtt}
\usepackage[all]{xy}
\usepackage{MinionPro}

\begin{document}
\small


\begin{center}
{\large\bf Alljoyn: Quick Reference}
\end{center}

\vspace*{1cm}

\tableofcontents

\pagebreak

\section{Introduction}
AllJoyn service frameworks are pre-defined set of (library) interfaces.



\section{About Feature}
\subsection{{\bfseries{}\texttt{org.alljoyn.About}} interface}
About inteface is a pre-defined interface (not user-defined interface),
which is expected to be implemented by all AllJoyn devices.
This interface will allow the app to \bb{advertise itself so other apps can
  discover it}. 

\subsubsection{Method: {\bf\texttt{GetAboutData}}}
  \begin{alltt}
  Variant<String>[] GetAboutData(s languageTag);
  \end{alltt}
Returns an array of variant string data. A variant is a ``tagged union'' data
which contains  a tag for identifying what kind of data it is. For example, it
can be a pair \verb+("AppId", "1019394")+
The fields are:
\bit
       \w \verb+AppId+
       \w \verb+DefaultLanguage+
       \w \verb+DeviceName+
       \w \verb+DeviceId+
       \w \verb+AppName+
       \w \verb+Manufacturer+
       \w \verb+ModelNumber+
       \w \verb+SupportedLanguagesn+
       \w \verb+Description+
       \w \verb+DateOfManufacture+
       \w \verb+SoftwarVersion+
       \w \verb+AJSoftwareVersion+
       \w \verb+HardwareVersion+
       \w \verb+SupportUrl+
       \eit

\subsubsection{Method: {\bf\texttt{GetObjectDescription}}}
\begin{verbatim}
  a(oas) GetObjectDescription()
\end{verbatim}
Returns the array of object paths
   and the list of supported interfaces provided by each object.

\subsubsection{Signal: {\bf\texttt{Announce}}}
  \bit
  \w \verb+Announce+: used to announce the application info and interfaces
   it supports
     \bit
     \w \verb+uint16 version+
     \w \verb+uint16 port+: session port, the app will listen on this port for
     incoming sessions
     \w \verb+{ObjectPath, String[]}[] objectDescription+:
     \w \verb+Variant<string>[] metaData+: fields as given in \verb+GetAboutData+
     \eit
  \eit

%\eit


\section{Onboarding Service Framework}
\subsection{{\bf\texttt{org.alljoyn.Onboarding}} interface}
\subsubsection{Property {\bf\texttt{Version}}}
\begin{verbatim}
  uint16 Version
\end{verbatim}
\subsubsection{Property {\bf\texttt{State}}}
\begin{verbatim}
  int16 State
\end{verbatim}
Configuration state.
\subsubsection{Property {\bf\texttt{LastError}}}
\begin{verbatim}
  uint16 LastError
\end{verbatim}

\subsubsection{Property {\bf\texttt{ConfigWiFi}}}
\begin{verbatim}
  int16 ConfigWiFi(s SSID, s passphrase, n authType)
\end{verbatim}
Sends the personal AP (access point) info to the onboardee.
\subsubsection{Property {\bf\texttt{Connect}}}
\begin{verbatim}
  void Connect();
\end{verbatim}
Tells the onboardee to connect to the personal AP.
\subsubsection{Property {\bf\texttt{Offboard}}}
\begin{verbatim}
  void Offboard();
\end{verbatim}
Tells the onboardee to disconnect from the
  personal AP, clear the personal AP configuration fields, and start the soft
  AP mode .
\subsubsection{Property {\bf\texttt{GetScanInfo}}}
\begin{verbatim}
  void GetScanInfo();
\end{verbatim}
Scan all the WiFI APs in the
  onboardee's proximity.

\subsubsection{Signal {\bf\texttt{ConnectionResult}}}
\begin{verbatim}
  ConnectionResult
\end{verbatim}
     \bit
     \w \verb+int16 resultcode+
     \eit


\section{Notification Service Framework}
\subsection{{\bf\texttt{org.alljoyn.Notification}} interface}
Used by AllJoyn applications to send events or state update information to
other devices connected to an end-user's home network (e.g. WiFi network).
\subsubsection{Property {\bf\texttt{Version}}}
\subsubsection{Signal {\bf\texttt{Notify}}}
     \bit
     \w \verb+short version+: version of the notification protocol
     \w \verb+integer msgId+: unique identification assigned to the
     notificatino message, assigned by the Notification Service Framework
     \w \verb+short msgType+: 0: emergency, 1: warning, 2: information
     \w {\bf\texttt{string deviceID}}: globally-unique id for the given device
                                        (instance) 
     \w \verb+string deviceName+: 
     \w {\bf{\texttt{byte[] appID}}}: globally-unqiue identififer (GUID) for a
                                    given AllJoyn application
     \w \verb+string appName+
     \w \verb+list<langText> langTexts+
     \w \verb+list<attribute> attributes+
     \w \verb+list<customAttribute> customAttributes+
     \eit
\subsection{{\bf\texttt{org.alljoyn.Notification.Producer}} interface}
Notification producer intefrace is announced s.t., when a device scans the
network, it can find all producer devices.
\subsubsection{Method {\bf\texttt{Dismiss}}}
\begin{verbatim}
  Dismiss(integer msgId);
\end{verbatim}
Notifies the producer that a notification
  has been dismissed.

\subsection{{\bf\texttt{org.alljoyn.Notification.Dismisser}} interface}
The Dismiss sessionliess signals are sent to notify other consumers on the
proximal network that a notification has been dismissed.

\subsubsection{Property {\bf\texttt{Version}}}

\subsubsection{Signal {\bf\texttt{Dismiss}}}
Notifies consumers that the
  noficiation has been dismissed.
    \bit
    \w \verb+uint16 msgId+
    \w \verb+byte[] appId+
    \eit

\section{Configuration Service Framework}
\subsection{{\bf\texttt{org.alljoyn.Config}} interface} 
This interface is for device-specific configuration.

\subsubsection{Property {\bf\texttt{Version}}}
\subsubsection{Method {\bf\texttt{FactoryReset}}}
\subsubsection{Method {\bf\texttt{Restart}}}
\subsubsection{Method {\bf\texttt{SetPasscode}}}
\subsubsection{Method {\bf\texttt{GetConfigurations}}}
\subsubsection{Method {\bf\texttt{UpdateConfigurations}}}
\subsubsection{Method {\bf\texttt{ResetConfigurations}}}
\subsubsection{Method {\bf\texttt{ResetConfigurations}}}



\end{document}
