\documentclass{myart}
\usepackage{mydef,epsfig}
\def\sbf{\sf\bfseries}
\def\Q{{\sf{}Q}}
\def\A{{\sf{}A}}
\begin{document}
\title{\large\bf Notes on Computer Architecture \vspace*{-0.5cm}}
\author{\normalsize Cheoljoo Jeong $\arc{\mbox{cjeong@cs.columbia.edu}}$}
\date{}
\maketitle
\small

\section{Hazards}
\subsection{Hazards in P/H basic 5-stage pipeline}
\subsubsection{Structural hazards}
\bit
\w \ee{happens when there exist resource contentions}
\w \bb{register file}: dual-ported read, single-ported write
  \bit
  \w \Q: what if we need two register writes in one cycle?
  \w \A: not in 5-stage pipeline. reg-write happens only in WB stage. and only
  one instruction can be in WB stage.
  \w \Q: if it happens in other pipelines?
  \w \A: dual-ported write register file. or all but one ``write'' must STALL.
  \eit
\eit
\subsection{Data hazards}
\bit
\w \ee{happens when pipeline changes order of an access to operands}
\w \bb{solutions}:
   \bit
   \w \bb{forwarding/bypassing}: FU (forwarding unit) 을 만들어서, ``어느 때''
   forwarding이 필요한지를 detect;
   \w FU는 \bb{producer instruction}은 무엇이고, {consumer instruction}은
   무엇인지 나열한 후, 각 pair에 대해 \bb{condition}을 기술하고, \bb{action}을
   기술함 
      \bit
      \w e.g. IR.EX/MEM[op] = R-type 이고 IR.ID/EX[op] = R-type 이고,
              IR.EX/MEM[rd] = IR.ID/EX[rs] 이면, ...
      \eit 
   \w Forwarding의 \bb{source points}:
       \ben
       \w ALU result
       \w Data Memory output
       \een
   \w Forwarding의 \bb{destination points}:
       \ben
       \w ALU input A 
       \w ALU input B
       \w Data memory (In)
       \w COND $\stackrel{?}{>} 0$
       \een
   \eit
\w \bb{load delay}: extra latency to produce load result (일반 ALU instruction
경우, EX stage에서 result, LOAD instr.의 경우, MEM stage에서 result) 
\w \bb{Classification}  
   \bit
   \w RAW (read after write)
   \w WAW (write after write): NO since write happens only in WB
   \w WAR (write after read): 
   \w RAR (read after read): NOT a hazard
   \eit
\w \Q: Memory location (instead of regs) 을 통한 data hazard는 가능한가?
\w \A: No. since MEM stage는 instruction-order를 preserve
\w \bb{HDU (hazard detection unit)}: forwarding에 의해 해결 불가능한 경우에,
   stall시켜주는 unit
   \ben
   \w freeze PC \& freeze IF/ID
   \w advance EX-WB forward
   \w inject no-op (bubble) to EX stage (즉, ID/EX pipeline reg)
   \een
\eit
\subsection{Control hazards (a.k.a. branch hazards)}
\bit
\w POLICY\#1: \bb{ALWAYS STALL} (3 cycle penalty)
\w POLICY\#1': \bb{ALWAYS STALL, Use ``optimized'' pipeline\footnote{BTA와 COND
  resolution을 ID stage에서}} (2 cycle penalty)
\w POLICY\#2A: \bb{Assume ``Branch Not Taken'' (= ``Branch unlikely'')}
   \bit
   \w Idea: fall-through does not stall
   \w Penalty: 실제로 branch NOT taken인 경우 penalty 0, branch taken인 경우,
   penalty 1 
   \eit
\w POLICY\#2B: \bb{Assume ``Branch Taken'' (= ``Branch likely'')}
   \bit
   \w Idea: branch does not stall
   \w Penalty: 실제로 branch taken인 경우 penalty 0, branch NOT taken인 경우,
   penalty 1 
   \eit
\w POLICY\#3: \bb{Branch delay slot (COMPILER solution)}
   \bit
   \w Idea: we always execute the instr. following branch; schedule something
   useful for this delay slot
   \w Variants: FROM FALL-THROUGH, FROM TARGET, FROM BEFORE
   \eit
\w POLICY\#4: \bb{Canceling/Nullifying branch}
\eit
\subsection{Hazards in ``Multicycle Operations''}
\bit
\w EX-units: 
   \bit
   \w integer unit (1 cycle): handles load/store (EA calculation), int ALU
   ops, branches
   \w FP add/sub (4 cycles): pipelined
   \w FP/int multiplier (7 cycles): pipelined
   \w FP/int divider (40 cycles): NOT pipelined
   \eit
\eit


\end{document}
