\documentclass{beamer}
\useinnertheme{rectangles}
\usepackage{graphicx}
\usepackage{mydef}
\usepackage{MinionPro}
\usefonttheme{serif}
\usepackage[all]{xy}

\title{Distributed Runtime}
\author{Cheoljoo Jeong}
\date{}
\begin{document}

\begin{frame}
\titlepage
\end{frame}


\begin{frame}
\frametitle{Distributed Runtime for The Meadow System}

The Meadow system consists of many {\em devices\/} which contains {\em
  processes\/} reactive to {\em events\/}. 

\vspace*{0.3cm}

Each device includes a \bb{runtime} which is reponsible for routing
events between devices and schedule the execution of processes upon
arrival of events.

\vspace*{0.3cm}

A process can have a {\em state\/} and the states of processes are maintained
by the runtime\footnote{Or should we assume that the process body is
  non-time-consuming?}.

\vspace*{0.3cm}

Runtimes collectively implements ``{\em{}global lookup tables\/}.''

\end{frame}


\begin{frame}[fragile]
\frametitle{Workflow Graphs}

A user provides {\em workflow definitions}, which will be compiled into
{\em workflow graphs\/}. 

\vspace*{0.4cm}

A \bb{workflow graph} is $G = (V, E)$, where $V$ is a set of 
reactive processes and $E$ is a set of directed, predicated edges between
processes. 

\vspace*{0.4cm}

The runtime in the device is reponsible for {\em execution of workflow
  graphs\/}, which involves 1) launching of processes and 2) routing of event
instances to other runtime.

\end{frame}

\begin{frame}[fragile]
\frametitle{Workflow Graphs (Cont.)}

A \bb{vertex} of a work process represents a process. A workflow process
performs one of the following activities:

\vspace*{0.3cm}

\bit
\w \bb{Computation}: This happens usually execution of the ``passive
callbacks'' by the runtime.

\vspace*{0.3cm}

\w \bb{Event instance generation}: An event instance is generated, which will
be routed by the runtime.

\eit

\vspace*{0.4cm}

A \bb{directed edge} of a workflow graph is $(u, v, C)$, where $u, v$ are
indicient vertices and $C$ is a condition over the events to which the graph
is sensitive to.

\end{frame}


%
% Implementation
%

%% \begin{frame}[fragile]
%% \frametitle{Q\#1: How to Specify Process Behavior}

%% \bb{\large CASE \#2: Provide a library}
%% \bit
%% \w Library which contains functions which interact with
%%   the daemon.
%% \w Each ``\verb+@(Event e)+'' must be implemented as an event loop. So, the
%%    library may need to provide some thread management facility.
%% \eit

%% \vspace*{0.3cm}


%% \end{frame}

\end{document}

%%  LocalWords:  RFID DSMS SQL CEP IFP Dataflow cep compsys softsys
